Visualizing probability distributions across bivariate cyclic temporal granularities

### 2020-03-05 main.Rmd
spell check

abstract: |
 Recent advances in technology greatly facilitates recording and storing data at much finer temporal scales than was previously possible. As the frequency of time-oriented data increases, the number of questions about the observed variable that can be addressed by visual representation also increases. We propose some new tools to explore this type of data, which deconstruct time in many different ways. There are several classes of time deconstructions including linear and cyclic time granularities. Linear time granularities respect the linear progression of time such as hours, days, weeks and monthsCyclic time granularities can be circular such as hour of the day, and day of the week, quasi-circular such as day of the month, and aperiodic such as public or school holidays.

@@

Linear time granularities respect the linear progression of time such as hours, days, weeks and months, and non-repeating. One of the first attempts to characterize these granularities is due to @Bettini1998-ed. However, the definitions and rules defined are inadequate for describing cyclic or repeating granularities. Hence, there is a need to define some new cyclic time granularities, that can be useful in visualizationsCyclic time granularities can be circular, quasi-circular or aperiodic. Examples of circular granularities are hour of the day and day of the week; an examples of a quasi-circular granularity is day of the month; examples of aperiodic granularities are public holidays and school holidays.

@@

cyclic time granularities with periodicities can be constructed by relating two linear granularities. The mappings between these linear granularities can be regular (minutes and hours) or irregular (days and months). <!--For example, a regular mapping exists between  (where 60 minutes always add up to 1 hour), while an irregular mapping exists between  (as the number of days per month ranges from 28 to 31).--> A grouping operation in Calendar algebra could lead to a regular mapping, whereas a altering-tick operation could result in an irregular mapping. Cyclic time granularities are referred to as **circular** if mappings are regular and **quasi-circular** if the mapping is irregular. Examples of common circular granularities include hour of the day, and day of the week, whereas examples for quasi-circular granularities can be day of the month or week of the month.

@@

# filter(fielding_wckts %in% c(0,1)) %>%
cricket_data %>% prob_plot("over", "lag_field",
  hierarchy_model,
  response = "run_rate",
  plot_type = "violin",
  quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9)
) + ggtitle("") + geom_boxplot(width = 0.5)
```

>>>>

abstract: |
 Recent advances in technology greatly facilitates recording and storing data at much finer temporal scales than was previously possible. As the frequency of time-oriented data increases, the number of questions about the observed variable that can be addressed by visual representation also increases. We propose some new tools to explore this type of data, which deconstruct time in many different ways. There are several classes of time deconstructions including linear and cyclic time granularities. Linear time granularities respect the linear progression of time such as hours, days, weeks and months. Cyclic time granularities can be circular such as hour of the day, and day of the week, quasi-circular such as day of the month, and aperiodic such as public or school holidays.

@@

Linear time granularities respect the linear progression of time such as hours, days, weeks and months, and non-repeating. One of the first attempts to characterize these granularities is due to @Bettini1998-ed. However, the definitions and rules defined are inadequate for describing cyclic or repeating granularities. Hence, there is a need to define some new cyclic time granularities, that can be useful in visualizations. Cyclic time granularities can be circular, quasi-circular or aperiodic. Examples of circular granularities are hour of the day and day of the week; an examples of a quasi-circular granularity is day of the month; examples of aperiodic granularities are public holidays and school holidays.

@@

Cyclic time granularities with periodicities can be constructed by relating two linear granularities. The mappings between these linear granularities can be regular (minutes and hours) or irregular (days and months). <!--For example, a regular mapping exists between  (where 60 minutes always add up to 1 hour), while an irregular mapping exists between  (as the number of days per month ranges from 28 to 31).--> A grouping operation in Calendar algebra could lead to a regular mapping, whereas a altering-tick operation could result in an irregular mapping. Cyclic time granularities are referred to as **circular** if mappings are regular and **quasi-circular** if the mapping is irregular. Examples of common circular granularities include hour of the day, and day of the week, whereas examples for quasi-circular granularities can be day of the month or week of the month.

@@

# filter(fielding_wckts %in% c(0,1)) %>%
cricket_data %>% prob_plot("over", "lag_field",
  hierarchy_model,
  response = "run_rate",
  plot_type = "violin",
) + ggtitle("") + geom_boxplot(width = 0.5, aes(colour = lag_field)) + ylim(0,4.5) +
      theme(
      axis.text = element_text(size = 24),
      strip.text = element_text(size = 24),
      axis.title = element_text(size = 24),
      legend.title = element_text(size = 24),
      legend.text = element_text(size = 24),
      legend.position = "bottom",
      plot.title =  element_text(size = 24)
    )
```
